An Answer Set Programming Framework for Reasoning about Agents' Beliefs and Truthfulness of Statements

### 2020-06-11 main.tex
Updates from Overleaf

@@

The set of observations $O_w\subseteq H^n$ in Example~\ref{ex1} includes the observations such as 
%`John comes from a poor family' at time point $t_0$, 
\begin{itemize}
    \item `John attends an expensive college' at time point $t_1$, and 
    \item `John receives a need-based scholarship' at time point $t_2$.
\end{itemize}
 In this particular example we do not have any  action occurrences, i.e., $O_a = \emptyset$.\footnote{Some frameworks for reasoning about actions and change may require the introduction of a NOP action for modeling the example.} Our default theory $Def$ consists of defaults $d_1$ and $d_2$ stated earlier, which allow us to make conclusions regarding our agent's beliefs.

\iffalse
%TEMP
\begin{figwindow}[1,r,%
        {\includegraphics[width=0.48\textwidth]{newstep.pdf}},%
        {Answering Question: $(H^n,Act,Def) \models p[t]$?\label{fig:schema} }]
\end{figwindow}
\fi
Let us consider a theory $T = (H^n, Act, \textit{Def})$ and the associated $\Phi_{Act}$ and $\models_D$. First, we are interested in defining the beliefs held by the agent over time. And, then, we are interested in using this information for answering the question of whether a statement $b[t]$,  made at time step $t$, where $b \in B$, can be accepted as truthful. For this, we define the entailment relation $\models$ between $T$ and $b[t]$. Intuitively, the process entails the steps below: %(Figure~\ref{fig:schema}):

@@

Let us start with a simpler case---where we do not worry about action
occurrences. We represents the components of the theory by means of keywords 
$stm$ (for $statement$), $obs$ (for $observed$), $rule$, $default$, and $prefer$. 
Fluent literals are defined as usual. Similarly, symbols for beliefs yield the notion of belief literal. By literals we mean fluent literals or belief literals. $\overline{l}$ 
denotes the complement of literal $l$. Agents can make statement about literals.  
Statements are given in the form {$stm(b,s)$}
where  $s$ is a non-negative integer (representing a time step) and $b$ is a belief literal.  
For example, the statement `John said that he is poor'  can be stated as 
$stm(poor, 0)$.

@@

Intuitively, $r$ and $d$ are used to name rules and defaults. 
%
%We will next define the notion of consistent knowledge bases. 
We say that a set of observations $O$ is \emph{consistent} if, for each time step $s$ and fluent literal $l$, $\{obs(l,s), obs(\neg l, s)\} \not\subseteq O$. 
Following the conventional assumption in the literature, we will make the following assumptions about the preference relation in $RD$: 
\emph{acyclic preference}: we assume that the preference relation between defaults represents a transitive and irreflexive relation, i.e., there exists no sequence of preference statements $prefer(d_i, d_{i+1}, body_i)$, $i=1,\ldots,k$, in $RD$ such that $\bigcup_{i=1}^k body_i$  is consistent.

@@

A set of rules $RD$ is \emph{consistent} if, for any consistent set of fluent literals $X$, the logic program 
$\{b \leftarrow body  \mid rule(r, b, body) \in RD\} \cup \{l \leftarrow \:\mid\: l\in X\}$
is consistent.
%
\begin{definition}
A $KB = \langle O, RD \rangle$ is \emph{consistent} if $O$ and $RD$ is consistent. 
\end{definition} 
%
Our goal is to identify the agent's beliefs and to use them to determine whether  a statement $stm(p, s)$ is true at a time step $t$ given $KB$. We will develop a program $\Pi(KB)$ to answer this question. We assume a finite horizon of steps, denoted by $s(0), \ldots, s(k)$. We use $h(L,S)$ to encode that fluent literal $L$ is true at step $S$. We use $believes(B,S)$ to represent that the agent believes $B$ at step $S$. Furthermore, to simplify the use of $\Pi(KB)$ with current ASP solvers, we  encode a rule of the form \eqref{rule}  using {\bf (a)} the ASP atom $rule(r, head, name)$ and
{\bf (b)} the set of ASP atoms $\{mbr(l, name) \mid l \in body_{r}\}$, where \(name\) is a unique identifier for the rule; similar encodings are used for  \eqref{default} and \eqref{pref}. 
This method of encoding is similar to what is used  in answer set planning and allows for the separation between  the encoding of the problem and the rules for reasoning with it.
The main challenge to be addressed lies in the interaction between observations and defaults. 
We apply the following principles:
%
\begin{itemize}
\item[]{\bf (O)}ptimistic: The most recent observation reflects the true state of the world;
\item[]{\bf (SK)}eptical: When conflicting conclusions can be drawn then do not believe in any.
\end{itemize}

@@

 For reasoning about statements of the form \eqref{rule}, we have:  %
%
\begin{equation}\label{lp:rule}
\begin{array}{ll}
believes(H, S) \leftarrow & s(S),  rule(D, H, M),  \\
        & N  = \# count \{L : mbr(L, M)\},\\
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.8in}h(L, T) \}= N_h,  \\ 
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.44in}believes(L, T) \}= N_b, \\
        & N_h+N_b=N.  \\ 
%
%& N = \# count \{L : mbr(L, M),  h(L, S) \},\\ 
%     & N  = \# count \{L : mbr(L, M)\}.
      \end{array}
\end{equation}

%
Intuitively, (\ref{lp:rule}) states that, if the body of a $rule$ statement is satisfied, then its head must be true. Recall that, according to the syntax given earlier, the head is a belief -- hence the expression $believes(H,S)$ in (\ref{lp:rule}). Checking whether the body is satisfied is compactly achieved by leveraging the $\#count$ aggregate of the ASP Core 2 language\footnote{Details can be found in \cite{ASPCore2Standard}}, which calculates the cardinality of a set. For instance, the expression $\# count \{L : mbr(L, M), h(L, T) \}$ finds number of literals that are in the body of the rule ($mbr(L,M)$) and hold at time step $T$ ($h(L,T)$). The combination of aggregates present in (\ref{lp:rule}) checks that the number of literals that hold or are believed equals the size of the body of the rule.\footnote{Recall that fluent symbols and belief symbols are disjoint. }

 For reasoning about defaults  of the form \eqref{default} we have:
% 
%%\iffalse
%{\small
\begin{eqnarray}
app(D, H, S) & \leftarrow  &  s(S), default(D, H, M), \nonumber \\
        & & N  = \# count \{L : mbr(L, M)\},    \nonumber\\
        & & \# count \{L : mbr(L, M),  \nonumber\\
        & & \hspace*{.8in}h(L, T) \}= N_h,  \nonumber\\ 
        & & \# count \{L : mbr(L, M), \nonumber\\
        & & N_h+N_b=N. \label{lp:app} \\ 
%     & & NC  = \# count \{L : mbr(L, M)\}, \nonumber \\ 
%    & & \# count \{L : mbr(L, M),     h(L, T) \} == NC. \nonumber \\ 
ab(D, H, S)& \leftarrow &   s(S), app(D, H, S), \nonumber\\
    & &  neg(H, H_1), app(D_1, H_1, S),  \nonumber \\
    &&   \naf defeated(D_1, H_1, S).  \label{lp:ab}      \\
defeated(D, H, S) & \leftarrow  & s(S), app(D, H, S), \nonumber\\
        & &  neg(H,H_1), \nonumber \\
        & &  believes(H_1, S).  \:\:\:  \label{lp:defeat0} \\ 
defeated(D, H, S) & \leftarrow  & s(S),  app(D, H, S),    \nonumber \\ 
        & &  neg(H,H_1), \nonumber \\ %
        & & app(D_1, H_1, S), \nonumber\\
        & & prefer(D_1, D, S).   \label{lp:defeat}\\
believes(H, S) & \leftarrow & app(D, H, S), \nonumber \\
        & & \naf ab(D, H, S), \nonumber\\
        & & \naf defeated(D, H, S).  \label{lp:holds}
%    
%defeated(D, H, S)& \leftarrow &  s(S), app(D, H, S), h(\overline{H}, S).\label{lp:defeat0} \\
%    &&  app(D, H, S),  \nonumber \\
%    &&  h(\overline{H}, S). \nonumber \\       
%defeated(D, H, S)& \leftarrow  & s(S),  app(D, H, S), \label{lp:defeat} \\
%    && app(D_1, \overline{H}, S), prefer(D_1, D, S).  \nonumber \\
%    &&  prefer(D_1, D, S).  \nonumber      \\
%h(l_1, S), \ldots, h(l_k, S), \naf ab(d, l, S)   \label{lp:def}\\
%b(l, S) \leftarrow b(l_1, S), \ldots, b(l_k, S), \naf ab(d, l, S)   \label{lp:def_bel}\\
%app(d, l, S) \leftarrow h(l_1, S), \ldots, h(l_k, S) 
%h(H, S)& \leftarrow  app(D, H, S), \naf ab(D, H, S),  \label{lp:holds} \\
% \naf ab(D, H, S), \nonumber \\ 
% & \naf defeated(D, H, S). \nonumber 
\end{eqnarray} 
%} 
%%\fi
%
\iffalse
{
        \[
        \begin{array}{llr}
     app(D, H, S)  \leftarrow &  s(S), default(D, H, M),  \nonumber\\
        & N  = \# count \{L : mbr(L, M)\},  & \label{lp:app} \\
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.8in}h(L, T) \}= N_h,  \\ 
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.44in}believes(L, T) \}= N_b, \\
        & N_h+N_b=N.  \\ 
        ab(D, H, S) \leftarrow  & s(S), app(D, H, S), neg(H,H_1), & (10)\\   %\label{lp:ab}
        &  app(D_1, H_1, S), \\
        & \naf defeated(D_1, H_1, S).        \\
        defeated(D, H, S) \leftarrow  & s(S), app(D, H, S), neg(H,H_1), \\
        & believes(H_1, S).  \:\:\: & (11) \\ %\label{lp:defeat0} 
        defeated(D, H, S) \leftarrow  & s(S),  app(D, H, S), neg(H,H_1), & (12) \\  %\label{lp:defeat}
        & app(D_1, H_1, S), \\
        & prefer(D_1, D, S).   \\
        believes(H, S) \leftarrow & app(D, H, S), \naf ab(D, H, S),\\
        & \naf defeated(D, H, S).  & (13) \\ %\label{lp:holds}
        \end{array}
        \]
        }
%
\fi

@@

 For reasoning about preferences of the form \eqref{pref}, we employ the rules:
% 
{
\begin{equation}\label{lp:pref}
\begin{array}{ll}
prefer(D_1, D_2, S) \leftarrow &  s(S),   prefer(D_1, D_2, M),   \\
     & N  = \# count \{L : \\
     &  \hspace*{.3in} mbr(L, M)\},  \\ 
     & \# count \{L : mbr(L, M),  \\
     &  \hspace*{.3in}  h(L, S) \} {=} N_h, \\
     & \# count \{L : mbr(L, M), \\
     & \hspace*{.3in} believes(L, S) \} {=} N_b, \\
     & N_h+N_b=N.
      \end{array}
\end{equation} 
}
%
This rule defines when a preference among two defaults can be applied. 
%
%\iffalse 
%Rule~\eqref{lp:app} defines when a default is \emph{applicable}, i.e., when its body is satisfied. The interaction between defaults and rules in the knowledge base are dealt with using rules~\eqref{lp:ab}---\eqref{lp:defeat}. Rule~\eqref{lp:ab} enforces the principle {\bf (SK)}. It says $ab(d, h, s)$ that if there is another default with the conflicting conclusion ($\overline{h}$) 
%which is not defeated then $ab(d, h, s)$ is true. This atom will be used to suppress the conclusion of $d$.   
%\eqref{lp:defeat0} dismisses the applicability of a default if the contrary of its conclusion is already established. On the other hand, rule \eqref{lp:defeat} expresses that a default is defeated if there is a more preferred default with conflicting conclusion that is applicable.  
%
%
% 
%
%\item Rules for reasoning about beliefs  
%\begin{align}
%b(L, S) \leftarrow h(L, S)  \label{lp:belief}   
%\end{align} 
%
%
%\item For interacting between defaults, we have the following rules: 
%
%%\memo{
%%}
%
%\begin{center} 
%  \begin{tikzpicture} 
%%    \tikzstyle{every node}=[draw,shape=circle} 
%%    \node (s1) at (0:0) {$\alert{s_3:}$}   
%%      \draw[fill=blue!20,style=dashed] (-1.0,-1.0) rectangle (3.0,3.0)  ;
%
%      \draw (0,0) circle(0.4) node {$\textcolor{red}{\small s{-}1}$} ;
%      \draw  (2,0) circle(0.4) node  {$\textcolor{red}{\small s}$} ;
%     \draw  (4,0) circle(0.4) node {$\textcolor{red}{\small s{+}1}$} ;  
% 
%     \draw[-,style=dashed] (-1.6,0) -- (-0.4,0 )  ; 
%     \draw[-] (0.4,0) -- (1.6,0)  ;  
%     \draw[-] (2.4,0) -- (3.6,0 )  ; 
%     \draw[-,style=dashed] (4.4,0) -- (5.6,0 )  ; 
% 
%  \end{tikzpicture} 
%\end{center} 
%
%
%\begin{align}\label{lp:ab} 
%ab(D_1, L, S) \leftarrow \: \: &     default(D_1, L, Body_1),   \\
%    &      app(D_1, L, S), \nonumber \\
%    &      default(D_2, \overline{L}, Body_2), \nonumber \\
%    &      app(D_2, \overline{L}, S), \nonumber \\
%    &      \naf app(D_2, S-1).  \nonumber
%\end{align} 
%
%This rule encodes the principle {\bf (P3)}. 
%
%\fi
%
%
%
%% We will assume that things behave in accordance to the law of inertia. So, we will add to the the program the rule 
%%
%% \begin{align}
%% h(l, S+1) \leftarrow h(l, S),  \naf h(\overline{l}, S)   \label{lp:inertial} \\
%% b(l, S+1) \leftarrow b(l, S),  \naf b(\overline{l}, S)   \label{lp:inertial_bel} 
%% \end{align} 
%
%

@@

%
\begin{proposition}
For each consistent $KB$,  there exists no answer set of $\Pi(KB)$ that contains $believes(l, t)$ and $believes(\overline{l},t)$ for some 
literal $l$ and time step $t$. 
\end{proposition} 
%
\noindent
{\bf Proof.}
The proof of this proposition relies on the fact that $believes(l, t)$ is in an answer set $A$ if the logic program  rule \eqref{lp:rule} (or \eqref{lp:holds}) is satisfied by $A$. If a rule \eqref{lp:rule} is satisfied by $A$ then a $KB$ contains a rule $rule(r,l,body)$ and $body$ is satisfied by $A$. Because of the consistency of $KB$, we know that there exists no rule $rule(r', \overline{l}, body')$ such that $body'$ is satisfied by $A$. For otherwise, $KB$ is inconsistent. This will imply that for every default $default(d, \overline{l}, body')$, if the default is applicable then the body of the rule \eqref{lp:defeat0} is satisfied and therefore $defeated(d)$ belongs to $A$. Consequently, the rule \eqref{lp:holds} is not satisfied for the default $d$, and hence, $believes(\overline{l}, t)$ cannot belong to $A$.

@@

\begin{proposition} 
For $\mathtt{x} \in \{\mathtt{t, b, u}\}$,
if  $\Pi_Q \cup \{\leftarrow  \mathtt{x}(b,t).\}$ does not have an answer set and  
$\Pi_Q \cup \{\leftarrow \naf \mathtt{x}(b,t)\}$   has an answer set then   $\Pi_Q \models \mathtt{x}(b, t)$. 
\end{proposition}  
% 
\noindent
{\bf Proof.}
Observe that $\Pi_Q \cup \{\leftarrow \mathtt{x}(b,t).\}$ does not have an answer set 
implies that $\Pi_Q$ either has no answer set or every answer set of $\Pi_Q$
contains $\mathtt{x}(b,t)$. The second condition implies that $\Pi_Q$ has at least one answer set. 
This implies that  $\Pi_Q \models \mathtt{x}(b, t)$.  
\hfill\qedsymbol{}

@@

In this section, we shift our focus to set $A$ of actions from the action signature $\langle F, B, A \rangle$. Following common practice in action languages \cite{GelfondL98}, we assume that each action is  associated with a set of literals, called its \emph{effects}, with each effect potentially predicated on a different set of literals, called \emph{preconditions}. A third set of literals, the \emph{executability conditions}, states when the action can be executed vs impossible. This information is encoded in statements of the form: 
$exec(a,body)$ and $causes(a,p,body)$ 
%\begin{align}
% exec(a,body) \label{exec} \\
% causes(a,p,body) \label{effect} 
%\end{align}
%
where $a \in A$ is an action, $p$ is a fluent literal, and $body$ is a set of fluent literals. 
 The first statement gives the executability conditions of action $a$, while the second states that, when $body$ holds, $p$ is one effect of $a$. We represent an \emph{action occurrence observation} (or {\em action occurrence}) by a statement of the form
 $occ(a,s)$
%\begin{equation}\label{occu}
%occ(a,s)
%\end{equation}
where $a\in A$ and $s$ is a time step. Recall that the notion of knowledge base already supports providing action occurrence observations, as  $H^{n}$ is a set of observations and action occurrences. The semantics of the above statements is encoded by suitable rules added to $\Pi(KB)$. 
 Note that the body of $exec$ and $causes$ statements is encoded by means of the $mbr$ predicate, as we did for rules and defaults in the previous section. 
We add to $\Pi(KB)$ the following rules:
 %and can be encoded   as follows. 
%
\begin{eqnarray}
obs(P, S)  & \leftarrow & s(S), occ(A, S), exec(A, M), \label{lp:exec} \\
& & mbr(P, M).  \nonumber    \\ 
obs(L, S+1) & \leftarrow & s(S),     occ(A, S), causes(A, L, M),\\
& &  N  = \# count \{X : mbr(X, M)\},  \label{lp:effect} \nonumber\\
& & \# count \{X : mbr(X, M),  h(X, T) \} = N  \nonumber
\end{eqnarray}

>>>>

%Marcello Balduccini, Michael Gelfond, Enrico Pontelli, Tran Cao Son: An Answer Set Programming Framework for Reasoning about Agents' Beliefs and Truthfulness of Statements

@@

The set of observations $O_w\subseteq H^n$ in Example~\ref{ex1} includes the observations such as 
%`John comes from a poor family' at time point $t_0$, 
\begin{itemize}
    \item `John attends an expensive college' at time point $t_1$, and 
    \item `John receives a need-based scholarship' at time point $t_2$.
\end{itemize}
 In this particular example we do not have any  action occurrences.\footnote{Some frameworks for reasoning about actions and change may require the introduction of a NOP action for modeling the example.} Our default theory $Def$ consists of defaults $d_1$ and $d_2$ stated earlier, which allow us to draw conclusions regarding our agent's beliefs.

\iffalse
%TEMP
\begin{figwindow}[1,r,%
        {\includegraphics[width=0.48\textwidth]{newstep.pdf}},%
        {Answering Question: $(H^n,Act,Def) \models p[t]$?\label{fig:schema} }]
\end{figwindow}
\fi
Let us consider a theory $T = (H^n, Act, \textit{Def})$ and the associated $\Phi_{Act}$ and $\models_D$. First, we are interested in defining the beliefs held by the agent over time. And, then, we are interested in using this information for answering the question of whether a statement $b[t]$,  where $b \in B$ and $t$ is the time step at which the statement was made, can be accepted as truthful. For this, we define the entailment relation $\models$ between $T$ and $b[t]$. Intuitively, the process entails the steps below: %(Figure~\ref{fig:schema}):

@@

Let us start with a simpler case---where we do not worry about action
occurrences. We represents the components of the theory by means of keywords 
$stm$ (for $statement$), $obs$ (for $observed$), $rule$, $default$, and $prefer$. 
Fluent literals are defined as usual. Similarly, symbols for beliefs yield the notion of belief literal. By literals we mean fluent literals or belief literals. $\overline{l}$ 
denotes the complement of literal $l$. Agents can make statements about belief literals.  
Statements are given in the form {$stm(b,s)$}
where  $s$ is a non-negative integer (representing a time step) and $b$ is a belief literal.  
For example, the statement `John said that he is poor'  can be stated as 
$stm(poor, 0)$.

@@

Intuitively, $r$ and $d$ are used to name rules and defaults. 
%
%We will next define the notion of consistent knowledge bases. 
We say that a set of observations $O$ is \emph{consistent} if, for each time step $s$ and fluent literal $l$, $\{obs(l,s), obs(\neg l, s)\} \not\subseteq O$. 
Following the conventional approach from the literature, we make the assumptions that the preference relation in $RD$ is an 
\emph{acyclic preference} and  that it is a transitive and irreflexive relation, i.e., there exists no sequence of preference statements $prefer(d_i, d_{i+1}, body_i)$, $i=1,\ldots,k$, in $RD$ such that $d_1 = d_{k+1}$ and $\bigcup_{i=1}^k body_i$  is consistent.

@@

A set of rules $RD$ is \emph{consistent} if, for any consistent set of fluent literals $X$, the logic program 
$\{b \leftarrow body  \mid rule(r, b, body) \in RD\} \cup \{l \leftarrow \:\mid\: l\in X\}$
is consistent.
%
\begin{definition}
A $KB = \langle O, RD \rangle$ is \emph{consistent} if $O$ and $RD$ is consistent. 
\end{definition} 
%
Our goal is to identify the agent's beliefs and to use them to determine whether  a statement $stm(p, s)$ is true at a time step $t$ given $KB$. We will develop a program $\Pi(KB)$ to answer this question. We assume a finite horizon of steps, denoted by $s(0), \ldots, s(k)$. We use $h(L,S)$ to encode that fluent literal $L$ is true at step $S$. We use $believes(B,S)$ to state that the agent believes $B$ at step $S$. Furthermore, to simplify the use of $\Pi(KB)$ with current ASP solvers, we  encode a rule of the form \eqref{rule}  using {\bf (a)} the ASP atom $rule(r, head, name)$ and
{\bf (b)} the set of ASP atoms $\{mbr(l, name) \mid l \in body_{r}\}$, where \(name\) is a unique identifier for the rule; similar encodings are used for  \eqref{default} and \eqref{pref}. 
This method of encoding is similar to what is used  in answer set planning and allows for the separation between  the encoding of the problem and the rules for reasoning with it.
The main challenge to be addressed lies in the interaction between observations and defaults. 
We apply the following principles:
%
\begin{itemize}
\item[]{\bf (O)}ptimistic: The most recent observation reflects the true state of the world;
\item[]{\bf (SK)}eptical: When conflicting conclusions can be drawn then do not believe in any.
\end{itemize}

@@

 For reasoning about statements of the form \eqref{rule}, we have:  %
%
\begin{equation}\label{lp:rule}
\begin{array}{ll}
believes(H, S) \leftarrow & s(S),  rule(D, H, M),  \\
        & N  == \# count \{L : mbr(L, M)\},\\
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.8in}h(L, T) \}== N_h,  \\ 
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.44in}believes(L, T) \}== N_b, \\
        & N_h+N_b==N.  \\ 
%
%& N = \# count \{L : mbr(L, M),  h(L, S) \},\\ 
%     & N  = \# count \{L : mbr(L, M)\}.
      \end{array}
\end{equation}

%
Intuitively, (\ref{lp:rule}) states that, if the body of a $rule$ statement is satisfied, then its head must be true. Recall that, according to the syntax given earlier, the head is a belief -- hence the expression $believes(H,S)$ in (\ref{lp:rule}). The `$==$' symbol denotes identity at the level of ground terms.\footnote{The variables involved are replaced by a constant during the grounding process and the identity is checked during the solving phase.} Checking whether the body is satisfied is compactly achieved by leveraging the $\#count$ aggregate of the ASP Core 2 language\footnote{Details can be found in \cite{ASPCore2Standard}}, which calculates the cardinality of a set. For instance, the expression $\# count \{L : mbr(L, M), h(L, T) \}$ finds number of literals that are in the body of the rule ($mbr(L,M)$) and hold at time step $T$ ($h(L,T)$). The combination of aggregates present in (\ref{lp:rule}) checks that the number of literals that hold or are believed equals the size of the body of the rule.\footnote{Recall that fluent symbols and belief symbols are disjoint. }

 For reasoning about defaults  of the form \eqref{default} we have:
% 
%%\iffalse
%{\small
\begin{eqnarray}
app(D, H, S) & \leftarrow  &  s(S), default(D, H, M), \nonumber \\
        & & N  == \# count \{L : mbr(L, M)\},    \nonumber\\
        & & \# count \{L : mbr(L, M),  \nonumber\\
        & & \hspace*{.8in}h(L, T) \}== N_h,  \nonumber\\ 
        & & \# count \{L : mbr(L, M), \nonumber\\
        & & N_h+N_b==N. \label{lp:app} \\ 
%     & & NC  = \# count \{L : mbr(L, M)\}, \nonumber \\ 
%    & & \# count \{L : mbr(L, M),     h(L, T) \} == NC. \nonumber \\ 
ab(D, H, S)& \leftarrow &   s(S), app(D, H, S), \nonumber\\
    & &  neg(H, H_1), app(D_1, H_1, S),  \nonumber \\
    &&   \naf defeated(D_1, H_1, S).  \label{lp:ab}      \\
defeated(D, H, S) & \leftarrow  & s(S), app(D, H, S), \nonumber\\
        & &  neg(H,H_1), \nonumber \\
        & &  believes(H_1, S).  \:\:\:  \label{lp:defeat0} \\ 
defeated(D, H, S) & \leftarrow  & s(S),  app(D, H, S),    \nonumber \\ 
        & &  neg(H,H_1), \nonumber \\ %
        & & app(D_1, H_1, S), \nonumber\\
        & & prefer(D_1, D, S).   \label{lp:defeat}\\
believes(H, S) & \leftarrow & app(D, H, S), \nonumber \\
        & & \naf ab(D, H, S), \nonumber\\
        & & \naf defeated(D, H, S).  \label{lp:holds}
%    
%defeated(D, H, S)& \leftarrow &  s(S), app(D, H, S), h(\overline{H}, S).\label{lp:defeat0} \\
%    &&  app(D, H, S),  \nonumber \\
%    &&  h(\overline{H}, S). \nonumber \\       
%defeated(D, H, S)& \leftarrow  & s(S),  app(D, H, S), \label{lp:defeat} \\
%    && app(D_1, \overline{H}, S), prefer(D_1, D, S).  \nonumber \\
%    &&  prefer(D_1, D, S).  \nonumber      \\
%h(l_1, S), \ldots, h(l_k, S), \naf ab(d, l, S)   \label{lp:def}\\
%b(l, S) \leftarrow b(l_1, S), \ldots, b(l_k, S), \naf ab(d, l, S)   \label{lp:def_bel}\\
%app(d, l, S) \leftarrow h(l_1, S), \ldots, h(l_k, S) 
%h(H, S)& \leftarrow  app(D, H, S), \naf ab(D, H, S),  \label{lp:holds} \\
% \naf ab(D, H, S), \nonumber \\ 
% & \naf defeated(D, H, S). \nonumber 
\end{eqnarray} 
%} 
%%\fi
%
\iffalse
{
        \[
        \begin{array}{llr}
     app(D, H, S)  \leftarrow &  s(S), default(D, H, M),  \nonumber\\
        & N  = \# count \{L : mbr(L, M)\},  & \label{lp:app} \\
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.8in}h(L, T) \}= N_h,  \\ 
        & \# count \{L : mbr(L, M),\\
        &\hspace*{.44in}believes(L, T) \}= N_b, \\
        & N_h+N_b=N.  \\ 
        ab(D, H, S) \leftarrow  & s(S), app(D, H, S), neg(H,H_1), & (10)\\   %\label{lp:ab}
        &  app(D_1, H_1, S), \\
        & \naf defeated(D_1, H_1, S).        \\
        defeated(D, H, S) \leftarrow  & s(S), app(D, H, S), neg(H,H_1), \\
        & believes(H_1, S).  \:\:\: & (11) \\ %\label{lp:defeat0} 
        defeated(D, H, S) \leftarrow  & s(S),  app(D, H, S), neg(H,H_1), & (12) \\  %\label{lp:defeat}
        & app(D_1, H_1, S), \\
        & prefer(D_1, D, S).   \\
        believes(H, S) \leftarrow & app(D, H, S), \naf ab(D, H, S),\\
        & \naf defeated(D, H, S).  & (13) \\ %\label{lp:holds}
        \end{array}
        \]
        }
%
\fi

@@

 For reasoning about preferences of the form \eqref{pref}, we employ the rules:
% 
{
\begin{equation}\label{lp:pref}
\begin{array}{ll}
prefer(D_1, D_2, S) \leftarrow &  s(S),   prefer(D_1, D_2, M),   \\
     & N  = =\# count \{L : \\
     &  \hspace*{.3in} mbr(L, M)\},  \\ 
     & \# count \{L : mbr(L, M),  \\
     &  \hspace*{.3in}  h(L, S) \} {=} =N_h, \\
     & \# count \{L : mbr(L, M), \\
     & \hspace*{.3in} believes(L, S) \} {=} =N_b, \\
     & N_h+N_b==N.
      \end{array}
\end{equation} 
}
%
This rule defines when a preference among two defaults can be applied. 
%
%\iffalse 
%Rule~\eqref{lp:app} defines when a default is \emph{applicable}, i.e., when its body is satisfied. The interaction between defaults and rules in the knowledge base are dealt with using rules~\eqref{lp:ab}---\eqref{lp:defeat}. Rule~\eqref{lp:ab} enforces the principle {\bf (SK)}. It says $ab(d, h, s)$ that if there is another default with the conflicting conclusion ($\overline{h}$) 
%which is not defeated then $ab(d, h, s)$ is true. This atom will be used to suppress the conclusion of $d$.   
%\eqref{lp:defeat0} dismisses the applicability of a default if the contrary of its conclusion is already established. On the other hand, rule \eqref{lp:defeat} expresses that a default is defeated if there is a more preferred default with conflicting conclusion that is applicable.  
%
%
% 
%
%\item Rules for reasoning about beliefs  
%\begin{align}
%b(L, S) \leftarrow h(L, S)  \label{lp:belief}   
%\end{align} 
%
%
%\item For interacting between defaults, we have the following rules: 
%
%%\memo{
%%}
%
%\begin{center} 
%  \begin{tikzpicture} 
%%    \tikzstyle{every node}=[draw,shape=circle} 
%%    \node (s1) at (0:0) {$\alert{s_3:}$}   
%%      \draw[fill=blue!20,style=dashed] (-1.0,-1.0) rectangle (3.0,3.0)  ;
%
%      \draw (0,0) circle(0.4) node {$\textcolor{red}{\small s{-}1}$} ;
%      \draw  (2,0) circle(0.4) node  {$\textcolor{red}{\small s}$} ;
%     \draw  (4,0) circle(0.4) node {$\textcolor{red}{\small s{+}1}$} ;  
% 
%     \draw[-,style=dashed] (-1.6,0) -- (-0.4,0 )  ; 
%     \draw[-] (0.4,0) -- (1.6,0)  ;  
%     \draw[-] (2.4,0) -- (3.6,0 )  ; 
%     \draw[-,style=dashed] (4.4,0) -- (5.6,0 )  ; 
% 
%  \end{tikzpicture} 
%\end{center} 
%
%
%\begin{align}\label{lp:ab} 
%ab(D_1, L, S) \leftarrow \: \: &     default(D_1, L, Body_1),   \\
%    &      app(D_1, L, S), \nonumber \\
%    &      default(D_2, \overline{L}, Body_2), \nonumber \\
%    &      app(D_2, \overline{L}, S), \nonumber \\
%    &      \naf app(D_2, S-1).  \nonumber
%\end{align} 
%
%This rule encodes the principle {\bf (P3)}. 
%
%\fi
%
%
%
%% We will assume that things behave in accordance to the law of inertia. So, we will add to the the program the rule 
%%
%% \begin{align}
%% h(l, S+1) \leftarrow h(l, S),  \naf h(\overline{l}, S)   \label{lp:inertial} \\
%% b(l, S+1) \leftarrow b(l, S),  \naf b(\overline{l}, S)   \label{lp:inertial_bel} 
%% \end{align} 
%
%

@@

%
\begin{proposition}
For each consistent $KB$,  there exists no answer set of $\Pi(KB)$ that contains $believes(l, t)$ and $believes(\overline{l},t)$ for some 
literal $l$ and time step $t$. 
\end{proposition} 
%
\noindent
{\bf Proof.}
The proof of this proposition relies on the fact that $believes(l, t)$ is in an answer set $A$ if the logic program  rule \eqref{lp:rule} (or \eqref{lp:holds}) is satisfied by $A$. If a rule \eqref{lp:rule} is satisfied by $A$ then a $KB$ contains a rule $rule(r,l,body)$ and $body$ is satisfied by $A$. Because of the consistency of $KB$, we know that there exists no rule $rule(r', \overline{l}, body')$ such that $body'$ is satisfied by $A$. For, otherwise, $KB$ is inconsistent. This will imply that for every default $default(d, \overline{l}, body')$, if the default is applicable, then the body of the rule \eqref{lp:defeat0} is satisfied and therefore $defeated(d)$ belongs to $A$. Consequently, \eqref{lp:holds} is not satisfied for the default $d$, and hence, $believes(\overline{l}, t)$ cannot belong to $A$.

@@

\begin{proposition} 
For $\mathtt{x} \in \{\mathtt{t, b, u}\}$,
if  $\Pi_Q \cup \{\leftarrow  \mathtt{x}(b,t).\}$ does not have an answer set and  
$\Pi_Q \cup \{\leftarrow \naf \mathtt{x}(b,t)\}$   has an answer set then   $\Pi_Q \models \mathtt{x}(b, t)$. 
\end{proposition}  
% 
\noindent
{\bf Proof.}
Observe that if $\Pi_Q \cup \{\leftarrow \mathtt{x}(b,t).\}$ does not have an answer set, this
implies that  either $\Pi_Q$ has no answer set or every answer set of $\Pi_Q$
contains $\mathtt{x}(b,t)$. The second condition implies that $\Pi_Q$ has at least one answer set. 
This implies that  $\Pi_Q \models \mathtt{x}(b, t)$.  
\hfill\qedsymbol{}

@@

In this section, we shift our focus to set $A$ of actions from the action signature $\langle F, B, A \rangle$. Following common practice in action languages \cite{GelfondL98}, we assume that each action is  associated with a set of literals, called its \emph{effects}, with each effect potentially predicated on a different set of literals, called \emph{preconditions}. A third set of literals, the \emph{executability conditions}, states when the action can be executed vs impossible. This information is encoded in statements of the form: 
$exec(a,body)$ and $causes(a,p,body)$ 
%\begin{align}
% exec(a,body) \label{exec} \\
% causes(a,p,body) \label{effect} 
%\end{align}
%
where $a \in A$ is an action, $p$ is a fluent literal, and $body$ is a set of fluent literals. 
 The first statement gives the executability conditions of action $a$, while the second states that, when $body$ holds, $p$ is one effect of $a$. We represent an \emph{action occurrence observation} (or {\em action occurrence}) by a statement of the form
 $occ(a,s)$
%\begin{equation}\label{occu}
%occ(a,s)
%\end{equation}
where $a\in A$ and $s$ is a time step. Recall that the notion of knowledge base already supports providing action occurrence observations, as  $H^{n}$ is a set of observations and action occurrences. The semantics of the above statements is encoded by suitable rules added to $\Pi(KB)$. 
 Note that the body of $exec$ and $causes$ statements is encoded by means of the $mbr$ predicate, as we did for rules and defaults in the previous section. 
We add to $\Pi(KB)$ the following rules:
 %and can be encoded   as follows. 
%
\begin{eqnarray}
obs(P, S)  & \leftarrow & s(S), occ(A, S), exec(A, M), \label{lp:exec} \\
& & mbr(P, M).  \nonumber    \\ 
obs(L, S+1) & \leftarrow & s(S),     occ(A, S), causes(A, L, M),\\
& &  N  = =\# count \{X : mbr(X, M)\},  \label{lp:effect} \nonumber\\
& & \# count \{X : mbr(X, M),  h(X, T) \} == N  \nonumber
\end{eqnarray}
